Validation of an Actuator Line Model Coupled to a Dynamic Stall Model for Pitching Motions Characteristic to Vertical Axis Turbines

### 2016-03-06 paper.tex
corrected spelling

    Vertical axis turbines (VATs) can be used to extract renewable energy from
    wind or water flows. A simpler design, low cost of maintenance, and the
    ability to accept flow from all directions perpendiculat to the rotor axis
    are some of the most important advantages over conventional horizontal axis
    turbines (HATs). However, VATs encounter complex and unsteady fluid
    dynamics, which present significant modeling challenges. One of the most
    relevant phenomena is dynamic stall, which is caused by the unsteady
    variation of angle of attack throughout the blade rotation, and is the focus
    of the present study. Dynamic stall is usually used as a passive control for
    VAT operating conditions, hence the importance of predicting its effects. In
    this study, \textbf{a coupled model is implemented with the open-source CFD
    toolbox OpenFOAM for solving the Navier--Stokes equations, where an actuator
    line model and dynamic stall model are used compute the blade loading and
    body force.} Force coefficients obtained from the model are validated with
    experimental data of pithing airfoil in similar operating conditions as an
    H-rotor type VAT. Numerical results show reasonable agreement with
    experimental data for pitching motion.

>>>>

    Vertical axis turbines (VATs) can be used to extract renewable energy from
    wind or water flows. A simpler design, low cost of maintenance, and the
    ability to accept flow from all directions perpendicular to the rotor axis
    are some of the most important advantages over conventional horizontal axis
    turbines (HATs). However, VATs encounter complex and unsteady fluid
    dynamics, which present significant modeling challenges. One of the most
    relevant phenomena is dynamic stall, which is caused by the unsteady
    variation of angle of attack throughout the blade rotation, and is the focus
    of the present study. Dynamic stall is usually used as a passive control for
    VAT operating conditions, hence the importance of predicting its effects. In
    this study, \textbf{a coupled model is implemented with the open-source CFD
    toolbox OpenFOAM for solving the Navier--Stokes equations, where an actuator
    line model and dynamic stall model are used compute the blade loading and
    body force.} Force coefficients obtained from the model are validated with
    experimental data of pithing airfoil in similar operating conditions as an
    H-rotor type VAT. Numerical results show reasonable agreement with
    experimental data for pitching motion.
